\bf Title here

### 2020-07-07 paper/paper.tex
make sure notation in the algorithm before matches with the ones in the global object

Now we formulate the problem in mathmetical notation. Projection pursuit
finds the projection basis
\(A^J = [\mathbf{a}_1, \cdots, \mathbf{a}_d]\), where \(\mathbf{a}_i\)
is a \(p \times 1\) vector, satisfying the following optimisation
problem:

\begin{align}
I &= \arg \max_{A \in \mathcal{A}^{p \times d}} f(A) \\
s.t. & \mathcal{A} = \{ \forall \mathbf{a}_i \in A: \lVert \mathbf{a}_i \rVert = 1 \}
\end{align}

via an iterative search: \(\{A^1, A^2, \cdots \cdots A^J\}\). Guided
tour interpolates geodesically between each basis found by projection
pursuit \(A^j\) to form a tour path:
\[\{A^1, A^1_1, A^1_2,\cdots, A^1_{k_1}, A^2, A^2_1, A^2_2, \cdots A^2_{k_2}, A^3, \cdots, A^{J-1}, A^{J-1}_{1}, \cdots A^{J-1}_{k_m} ,A^J  \}\]

@@

\citet{posse1995projection} used a stochastic direct search method, a
random search algorithm to sample a candidate bases \(A^{(j +1)}\) in
the neighbourhood of the current basis \(A^j\) defined by the radius of
the p-dimensional sphere, \(\alpha\). If the candidate basis has a
higher index value than the current basis, it is outputed as the target
basis \(A^{j + 1}\) along with other metadata and the algorithm stops.
If no basis is found to have higher index value after the maximum number
of tries \(i_{max}\), the algorithm stops with nothing outputted. \(h\)
is the halfing parameter meaning if \(h > 30\), the nieghbourhood
parameter \(\alpha\) will be reduced by half in the next iteration of
the random search. The algorithm for a random search is summarised in
Algorithm \ref{random-search}.

\newpage
\begin{algorithm}[H]
\SetAlgoLined
  initialisation: $h = 0$, $loop = 1, i_{max}$\;
  \While{$i < i_{max}$}{
    $A^{(j + 1)} = A^{j} + \alpha A_{rand}$\;
    $I^{(j + 1)} = f(A^{(j+1)})$\;
    \eIf{$I^{(j+1)} > I^{j}$}{
      $A^{j+1} = A^{(j+1)}$\;
      }{
      $h = h + 1$\;
      }
    $i = i + 1$\;
  }
  \caption{random search: my first algorithm}
  \label{random-search}
\end{algorithm}

@@

\newpage
\begin{algorithm}[H]
\SetAlgoLined
  initialisation: $loop = 1, i_{max}$\;
  \While{$i < i_{max}$}{
    Generate $\{A^{(j_1)}: A^{(1)}, \cdots A^{(5)}\}$\;
    Find $A^{(J_1)}$ where $J_1 \in j_1$ such that $\{A^{(J_1)}: \arg \max I(A^{(j_1)})\}$\;
    Denote $F_a = A^j$ and $F_z = A^{(J_1)}$\;
    Conduct singular value decomposition (SVD) for $F_a$ and $F_z$ such that $F_{a}^{T}F_z = V_a\Lambda V_z^T$\;
    Compute $G_a = F_aV_a$, $G_z = F_zV_z$ and $\tau = \arccos(\Lambda)$\;
    ...\;
    Find $A^{(J_2)}$ satisfying $A^{J_2}$ on the interpolation path such that $\{A^{(J_2)}: \arg \max I(A^{(j_2)})\}$\;
    Compute $I^{(J_2)} = f(A^{(J_2)})$, $p_{diff} = (I^{(J_2)} - I^{j})/I^j$\;
      \If{$p_{diff} > 0.001$}{
        $A^{j+1} = A^{(J_2)}$\;
        $I^{j+1} = I^{(J_2)}$\;
        \KwOut{$A^{j+1}$, $I^{j+1}$, $i$ and relevant metadata}
      }
    $i = i + 1$\;
  }
  \caption{searchgeodesic}
  \label{search_geodesic}
\end{algorithm}

Simulated annealing
\citep[\citet{kirkpatrick1983optimization}]{bertsimas1993simulated} is a
non-derivative procedure based on a non-increasing cooling scheme
\(T(i)\). Given an initial \(T_0\), the temperature at iteration \(i\)
is defined as \(T(i) = \frac{T_0}{log(i + 1)}\). The simulated annealing
algorithm works as follows. Given a neighbourhood parameter \(\alpha\)
and a randomly generated orthonormal basis \(A_{rand}\), a candidate
basis is constructed as
\(A^{(j+1)} = (1 - \alpha)A^{j} + \alpha A_{rand}\). If the index value
of the candidate basis is larger than the one of the current basis, the
candidate basis becomes the target basis. If it is smaller, the
candidate is accepted with probability
\(\text{prob} = \min\left\{\exp\left[\frac{I^j - I^{(j+1)}}{T(i)}\right],1\right\}\)
where \(I^j\) and \(I^{(j+1)}\) are the index value of the current and
candidate basis respectively. The algorithm can be summarised as in
Algorithm \ref{simulated_annealing}.

\newpage
\begin{algorithm}[H]
\SetAlgoLined
  initialisation: $T_0$, $\alpha$\;
  \While{$i < i_{max}$}{
    $A^{(j+1)} = (1 - \alpha)A^{j} + \alpha A_{rand}$ s.t. $\lVert\mathbf{a_i} \rVert = 1$\;
    $I^{(j+1)} = f(A^{(j+1)})$\;
    $T(i) = \frac{T_0}{\log(i + 1)}$\;
      \eIf{$I^{(j+1)} > I^{j}$}{
        $A^{j+1} = A^{(j+1)}$\;
        $I^{j+1} = I^{(j+1)}$\;
        \KwOut{$A^{j+1}$, $I^{j+1}$,$i$ and relevant metadata}
      }{
        $\text{prob} = \min\left\{\exp\left[\frac{I^j - I^{(j+1)}}{T(i)}\right],1\right\}$\;
        Draw $\text{rand} \sim \text{Unif(0, 1)}$\;
        \If{$\text{prob} > \text{rand}$}{
          \KwOut{$A^{j+1}$, $I^{j+1}$, $i$ and relevant metadata}
        }
      }
    $i = i + 1$\;
  }
  \caption{simulated annealing}
  \label{simulated_annealing}
\end{algorithm}

@@

\emph{Iterators} indexes the data collected and it has a time series
feature that prescribes the progression of the optimisation. Given the
complexity of the algorithm, there could be multiple iterators. The
simpliest iterator: \emph{index per observation}, is a unique identifier
for each observation in the data. For each level of iteration, we design
two indices: \emph{index per iteration} is fixed for each observations
in one iteration and has an increment of one once a new iteration
starts. \emph{index within iteration} has an increment of one for each
observation in an iteartion and starts over from one once a new
iteration starts. Depend on the number of iteration level (\(L\)), the
total number of iterators will differ and can be calculated as
\(1 + 2L\). In projection pursuit optimisation algorithms, there is only
one level of iteartion and hence exists three iterators: \(t_1\),
\(t_2\) and \(t_3\). A sketch of the design of iterators is provided in
Figure \ref{iterators}.

There could exist other parameters that are also of our interest but
can't be classified as one of the three categories. They are defined
under the fourth category: \emph{other parameter of interest: \(O\)}.
This could include variables that are specific for one particular
algorithm. Below we show an example of defining the global object of
projection pursuit guided tour.

Now we formulate the idea of global object in terms of mathmetical
notation. Superscript \(j\) is used to denote the number of iteration.
Candidate basis are denoted using subscript on the right and
interpolation basis are denoted using subscript on the left. The output
basis doesn't have subscript. For example, \(\mathbf{A}^1_{2}\) means
the second candidate basis in the first iteration and
\({}_{2}\mathbf{A}^1\) means the second interpolation basis that forms
the tour path from the current basis \(\mathbf{A}^0\) to the target
basis \(\mathbf{A}^1\). In the next iteration (iteration 2),
\(\mathbf{A}^1\) becomes the current basis, projection pursuit searches
through
\(\mathbf{A}^2_{1}, \mathbf{A}^2_{2}, \cdots, \mathbf{A}^2_{l_2}\) to
find the target basis \(\mathbf{A}^2\) and the tour path from
\(\mathbf{A}^1\) to \(\mathbf{A}^2\) is constructed using the
interpolation bases
\({}_{1}\mathbf{A}^2, {}_{2}\mathbf{A}^2, \cdots, {}_{k_2}\mathbf{A}^2\).
Below we arrange all the bases in a format that each coloumn represents
an iteration in Equation \ref{eq:matrixA}.

When aggregated with information of index value and other interesting
parameter, it is easier to write the basis as a column vector with its
natual ordering as in Equation \ref{eq:vectorA}. Hence the full data
structure can be written as in Equation \ref{eq:data-structure}

@@

@@

An optimisation procedure can easily generate hundreds or thousands
sampling points and the question becomes what are the points we want to
explore\ldots{} {[}the classification of searching or updating points
goes here and people should be clear what part of the points goes to the
plot - this is the data part not the plot part. {]}

In tour, all the points recorded in the global object can be divided
into two broad categories: searching points and interpolating points

\begin{itemize}
\item
  \emph{Searching points} include the observations recorded in the
  searching algorithm to find the target basis. The points for target
  bases is also included in the searching points and there is one such
  point per \texttt{tries}.
\item
  \emph{interpolating points} exist in the guided tour to produce
  continuous animated view from one target basis to another and it
  doesn't have \texttt{loop} value.
\end{itemize}

In the projection pursuit guided tour, the decision variable is the
projection matrix, the value of objective function is the index value.
There are three iterators: \texttt{id} is the smallest ordering unit
that increases by one for each observation; \texttt{tries} is updated
once a search-and-interpolate step is finished. A third iterator
\texttt{loop} is used to record the number of repetition in the search
step and starts over from one at the beginning of a new \texttt{tries}.
Three other parameters of our interest includes \texttt{method},
\texttt{alpha} and \texttt{info}. \texttt{method} identifies the name of
the searching method used and we are interested in comparing the
performance between different algorithms under direct search. The
neighbourhood parameter \texttt{alpha} controls the size of the sampling
space and we are interested to understand how the searching space
shrinks as the algorithm progresses. \texttt{info} labels different
stages in the searching process. A sketch of the global object for
projection pursuit guided tour is presented in Figure \ref{fig:glb-obj}.

\hypertarget{simulated-data}{%
\subsubsection{Simulated data}\label{simulated-data}}

>>>>

Now we formulate the problem in mathmetical notation. Projection pursuit
finds the projection basis
\(\mathbf{A}^J = [\mathbf{a}_1, \cdots, \mathbf{a}_d]\), where
\(\mathbf{a}_i\) is a \(p \times 1\) vector, satisfying the following
optimisation problem:

\begin{align}
I &= \arg \max_{\mathbf{A} \in \mathcal{A}^{p \times d}} f(\mathbf{A}) \\
s.t. & \mathcal{A} = \{ \forall \mathbf{a}_i \in \mathbf{A}: \lVert \mathbf{a}_i \rVert = 1 \}
\end{align}

via an iterative search:
\(\{\mathbf{A}^j: \mathbf{A}^1, \mathbf{A}^2, \cdots \cdots \mathbf{A}^J\}\).
Guided tour interpolates geodesically between each basis found by
projection pursuit \(A^j\) to form a tour path:
\[\{\mathbf{A}^1, {}_{1}\mathbf{A}^1, {}_{2}\mathbf{A}^1,\cdots, {}_{k_1}\mathbf{A}^1, A^2, {}_{1}\mathbf{A}^2, {}_{2}\mathbf{A}^2, \cdots {}_{k_2}\mathbf{A}^2, \cdots, A^{J-1}, {}_{1}A^{J-1}, \cdots {}_{k_m}A^{J-1} ,A^J \}\]

Here we use left subscript for the index of the interpolation at each
iteration i,e, \({}_{2}\mathbf{A}^1\) means the second interpolation
basis in the first iteration. The right subcript is reserved for the
index of searching points in the optimisation algorithm.

@@

\citet{posse1995projection} used a stochastic direct search method, a
random search algorithm to sample a candidate bases \(A^{j +1}_l\) in
the neighbourhood of the current basis \(A^j\) defined by the radius of
the p-dimensional sphere, \(\alpha\). If the candidate basis has a
higher index value than the current basis, it is outputed as the target
basis \(A^{j + 1}\) along with other metadata and the algorithm stops.
If no basis is found to have higher index value after the maximum number
of tries \(l_{\max}\), the algorithm stops with nothing outputted. \(c\)
is the halfing parameter meaning if \(c > 30\), the nieghbourhood
parameter \(\alpha\) will be reduced by half in the next iteration of
the random search. The algorithm for a random search is summarised in
Algorithm \ref{random-search}.

\newpage
\begin{algorithm}[H]
\SetAlgoLined
  initialisation: $c = 0$, $l = 1$ $l_{\max}$, $\mathbf{A}^{j}$\;
  \While{$l < l_{\max}$}{
    $\mathbf{A}^{j+1}_{l} = \mathbf{A}^{j} + \alpha \mathbf{A}_{rand}$\;
    $I^{j+1}_l = f(\mathbf{A}^{j+1}_l)$\;
    \eIf{$I^{j+1}_l > I^{j}$}{
      $\mathbf{A}^{j+1} = \mathbf{A}^{j+1}_{l}$\;
      }{
      $c = c + 1$\;
      }
    $l = l + 1$\;
  }
  \caption{random search: my first algorithm}
  \label{random-search}
\end{algorithm}

@@

\newpage
\begin{algorithm}[H]
\SetAlgoLined
  initialisation: $l_{\max}$, $\alpha$, $l_{\max}$,$\mathbf{A}^j$\;
  \While{$l < l_{\max}$}{
    Generate $\{A^{(j_1)}: A^{(1)}, \cdots A^{(5)}\}$\;
    Find $A^{J_1}_l$ where $J_1 \in j_1$ such that $\{A^{(J_1)}: \arg \max I(A^{(j_1)})\}$\;
    Denote $F_a = A^j$ and $F_z = A^{(J_1)}$\;
    Conduct singular value decomposition (SVD) for $F_a$ and $F_z$ such that $F_{a}^{T}F_z = V_a\Lambda V_z^T$\;
    Compute $G_a = F_aV_a$, $G_z = F_zV_z$ and $\tau = \arccos(\Lambda)$\;
    ...\;
    Find $A^{(J_2)}$ satisfying $A^{J_2}$ on the interpolation path such that $\{A^{(J_2)}: \arg \max I(A^{(j_2)})\}$\;
    Compute $I^{(J_2)} = f(A^{(J_2)})$, $p_{diff} = (I^{(J_2)} - I^{j})/I^j$\;
      \If{$p_{diff} > 0.001$}{
        $A^{j+1} = A^{(J_2)}$\;
        $I^{j+1} = I^{(J_2)}$\;
        \KwOut{$A^{j+1}$, $I^{j+1}$, $i$ and relevant metadata}
      }
    $i = i + 1$\;
  }
  \caption{searchgeodesic: this part has problem!}
  \label{search_geodesic}
\end{algorithm}

Simulated annealing
\citep[\citet{kirkpatrick1983optimization}]{bertsimas1993simulated} is a
non-derivative procedure based on a non-increasing cooling scheme
\(T(l)\). Given an initial \(T_0\), the temperature at iteration \(l\)
is defined as \(T(l) = \frac{T_0}{log(l + 1)}\). The simulated annealing
algorithm works as follows. Given a neighbourhood parameter \(\alpha\)
and a randomly generated orthonormal basis \(A_{rand}\), a candidate
basis is constructed as
\(A^{j+1}_l = (1 - \alpha)A^{j} + \alpha A_{rand}\). If the index value
of the candidate basis is larger than the one of the current basis, the
candidate basis becomes the target basis. If it is smaller, the
candidate is accepted with probability
\(\text{prob} = \\min\left\{\exp\left[\frac{I^j - I^{j+1}_l}{T(l)}\right],1\right\}\)
where \(I^j\) and \(I^{j+1}_l\) are the index value of the current and
candidate basis respectively. The algorithm can be summarised as in
Algorithm \ref{simulated_annealing}.

\newpage
\begin{algorithm}[H]
\SetAlgoLined
  initialisation: $T_0$, $\alpha$, $l_{\max}$,$\mathbf{A}^j$\;
  \While{$l < l_{\max}$}{
    $\mathbf{A}^{j+1}_{l} = (1 - \alpha)\mathbf{A}^{j} + \alpha \mathbf{A}_{rand}$ s.t. $\lVert\mathbf{a_i} \rVert = 1$\;
    $I^{j+1}_l = f(\mathbf{A}^{j+1}_{l})$\;
    $T(l) = \frac{T_0}{\log(l + 1)}$\;
      \eIf{$I^{j+1}_l > I^{j}$}{
        $\mathbf{A}^{j+1} = \mathbf{A}^{j+1}_{l}$\;
        $I^{j+1} = I^{j+1}_l$\;
        \KwOut{$\mathbf{A}^{j+1}$, $I^{j+1}$,$l$ and relevant metadata}
      }{
        $\text{prob} = \min\left\{\exp\left[\frac{I^j - I^{j+1}_l}{T(l)}\right],1\right\}$\;
        Draw $\text{rand} \sim \text{Unif(0, 1)}$\;
        \If{$\text{prob} > \text{rand}$}{
          \KwOut{$\mathbf{A}^{j+1}$, $I^{j+1}$, $l$ and relevant metadata}
        }
      }
    $l = l + 1$\;
  }
  \caption{simulated annealing}
  \label{simulated_annealing}
\end{algorithm}

@@

\emph{Iterators} indexes the data collected and has a time series
feature that prescribes the its natural ordering. Given the complexity
of the algorithm, there could be multiple iterators. The simpliest
iterator: \emph{index per observation}, is a unique identifier for each
observation in the data. For each level of iteration, we design two
indices: \emph{index per iteration} is fixed for each observations in
one iteration and has an increment of one once a new iteration starts.
\emph{index within iteration} has an increment of one for each
observation in an iteartion and starts over from one once a new
iteration starts. A sketch of the difference between these three
iterators is provided in Figure \ref{iterators}. Depend on the number of
iteration level (\(L\)), the total number of iterators will differ and
can be calculated as \(1 + 2L\).

@@

In projection pursuit optimisation algorithms, there is one level of
iteartion and hence exists three iterators: \texttt{id} indices each
observation by a unique number; \texttt{tries} is the index per
iteration iterator that gets updated once a search-and-interpolate step
is finished; and \texttt{loop} is the index within iteration iterator
and starts over from one at the beginning of a new \texttt{tries}.

There could exist other parameters that are also of our interest but
can't be classified as one of the three categories. They are defined
under the fourth category: \emph{other parameter of interest: \(O\)}. In
projection pursuit, three other parameters of our interest includes
\texttt{method}, \texttt{alpha} and \texttt{info}. \texttt{method}
identifies the name of the searching method used and we are interested
in comparing the performance between different algorithms under direct
search. The neighbourhood parameter \texttt{alpha} controls the size of
the sampling space and we are interested to understand how the searching
space shrinks as the algorithm progresses. \texttt{info} labels
different stages in the searching process. A sketch of the global object
for projection pursuit guided tour is presented in Figure
\ref{fig:glb-obj}.

\hypertarget{notation}{%
\subsection{Notation}\label{notation}}

Now we formulate the idea of global object using mathmetical notation.
Let \(\mathbf{A}^0\) denotes the starting basis. Superscript \(j\) is
used to denote the number of iteration (index per iteration). Candidate
basis are denoted using subscript on the right and interpolation basis
are denoted using subscript on the left. For example,
\(\mathbf{A}^1_{2}\) means the second candidate basis in the first
iteration and \({}_{2}\mathbf{A}^1\) means the second interpolation
basis that forms the tour path from the current basis \(\mathbf{A}^0\)
to the target basis \(\mathbf{A}^1\). In iteration two, \(\mathbf{A}^1\)
becomes the current basis and projection pursuit searches through
\(\mathbf{A}^2_{1}, \mathbf{A}^2_{2}, \cdots, \mathbf{A}^2_{l_2}\) to
find the target basis \(\mathbf{A}^2\). The tour path from
\(\mathbf{A}^1\) to \(\mathbf{A}^2\) is constructed using the
interpolation bases
\({}_{1}\mathbf{A}^2, {}_{2}\mathbf{A}^2, \cdots, {}_{k_2}\mathbf{A}^2\).
All the bases together can be arranged in a format that each coloumn
represents an iteration as in Equation \ref{eq:matrixA}.

@@

\hypertarget{simulated-data}{%
\subsection{Simulated data}\label{simulated-data}}
